Development of restored dike grasslands to reference needs more than four years and can hardly modified by local site conditions

### 2023-02-14 manuscript.Rmd
Update manuscript.Rmd

The vegetation developed in general in the right direction but was still distinct from positive references and seed mixtures after four years as it was shown for other seeding experiments or restorations [@mitchley2012; @Engst.2016; @Kaulfuss.2022] or for dike vegetation compared with semi-natural references [@Batori.2016]. This result is not surprising since the recovery debt is a general phenomenon of grassland restoration [@MorenoMateos.2017; @jones2018]. Our results support that dikes can promote biodiversity in agricultural landscapes [@Batori.2020] but the recovery debt highlights that restored grasslands cannot substitute old-growth grasslands [@nerlekar2020; @Buisson.2022].

@@

Variability important (different seed mixture types, different soils)

Higher seeding density not necessary for biodiversity

Outlook: landscape, management are important

>>>>

The vegetation developed in general in the right direction but was still distinct from positive references and seed mixtures after four years. It was rather similar to the negative reference of perennial ruderal vegetation. The gap between goal and restoration outcome was also shown for other seeding experiments or restorations [@mitchley2012; @Engst.2016; @Kaulfuss.2022] or for dike vegetation compared with semi-natural references [@Batori.2016]. This result is not surprising since the 'recovery debt' is a general phenomenon of grassland restoration [@MorenoMateos.2017; @jones2018].

@@

Our results suggest that adapted seed mixtures can increase restoration efficiency by seeding hay meadows on the north exposition and not on the south exposition of dikes. Furthermore, to reduce nutrient load by sand admixture has a clear positive effect, although the effect is small. The question remains if money is best invested in sand admixture or in other restoration measures. It seems not to be necessary to increase seeding density on dike slopes, which would reduce restoration costs. As well as the use of more deeper soil depth which would otherwise create costs for removal of topsoil. There was no perfect seed mixture-substrate combination, why we conclude that a variation of seed mixtures types and different soils along restoration sections would promote biodiversity better than a single solution [@bauer2023; @Holl.2022]. The strong and lasting negative effect of drought on the south exposition might require a resowing. To close the recovery debt, the adaption of the management might be promising since it is a relevant factor aside of restoration approach and site characteristics [@Grman.2013; @tölgyesi2021]. For example, starting sheep grazing on the restored sites will modify the disturbance regime and improve dispersal. All in all, our results support the finding that dike grasslands can promote biodiversity in agricultural landscapes [@Batori.2020]. Though, the recovery debt highlights that restored grasslands cannot substitute old-growth grasslands [@nerlekar2020; @Buisson.2022].
